---
title: "wordle" 
format: html
editor_options: 
  chunk_output_type: console
---

# Creating Wordle

##### Load required packages

```{r}
#| output: false
library(tidyverse)
```

### Step 1

###### Create a custom function called load_dictionary() that can be used to read in the data files

```{r}
load_dictionary <- function(filename) {
  words <- readLines(filename, warn = FALSE)
  words <- words[-1] # removes the header
  return(words)
}
```

###### Use the function to create two variables, solution_list and valid_list, that, respectively contain vectors of possible solution words and valid words to guess

```{r}
valid_list <- load_dictionary("collins-scrabble-words-2019.txt") # Official Scrabble words
solution_list <- load_dictionary("google-10000-english-usa-no-swears.txt") # Most common words in English language
# check whether it's correct
str(valid_list)
str(solution_list)
```

### Step 2

###### Winnow solution_list to only include words that are included in valid_list

```{r}
solution_list <- intersect(solution_list, valid_list) # Only include the common words that are official Scrabble words
```

There are 8,336 words in the updated vector.

### Step 3

###### Write a custom function called pick_solution() that \[1\] removes all words from solution_list that are not 5 letters in length, \[2\] randomly chooses a single word from those that remain, and \[3\] splits that word into a vector of single-character elements

```{r}
pick_solution <- function(vector, word_length) {
  five_letter_words <- vector[nchar(vector) == word_length] # only keeps the words of specified word_length
  solution_word <- sample(five_letter_words, size = 1) # picks one word at random, size gives the number of words to choose
  solution_characters <- str_split_1(solution_word, pattern = "") # splits the word into single character elements
  return(solution_characters)
}
solution <- pick_solution(solution_list, word_length = 5)
```

### Step 4

###### Create a function called play_wordle() that takes three arguments: [1] the answer to the puzzle, [2] a list of valid guesses, and [3] a value for "number of guesses."

###### It should [1] tell the player the rules of the game, [2] display what letters the player has not yet guessed, and [3] prompt the player for a guess, read in their guess, and check that their guess is valid.

Setting up a function to evaluate a player's guess
```{r}
evaluate_guess <- function(guess, answer) {
  feedback <- rep("-", length(solution)) # makes a blank feedback string

  for (letter in 1:length(answer)) { # for each letter in answer
    
    # correct letter in correct position
    if (guess[letter] == answer[letter]) {
      feedback[letter] <- "*"
      answer[letter] <- "-" # this removes the correct letter from the answer string
    }
    
    # correct letter in wrong position
    if (feedback[letter] == "-") {
    if (guess[letter] %in% answer) {
      feedback[letter] <- "+"
    }
    }
  }
  feedback
}
```

Setting up a function to play the game
```{r}
play_wordle <- function(answer, valid_guesses, num_guesses=6) {
  
  # print instructions
  print(paste0("You have ", num_guesses, " chances to guess a ", length(answer), "-letter word that has been selected at random from a list of possible words. ", "Each guess must be a valid ", length(answer), "-letter word."), quote = FALSE)
  print(paste0("*: letter is in the word and in the correct spot."), quote = FALSE) 
  print(paste0("+: letter is in the word but not in the correct spot."), quote = FALSE) 
  print(paste0("-: letter is not in the word."), quote = FALSE)
  starting_word <- rep("-", length(answer))
  print(paste(starting_word, collapse = " "))
  
  # setting the beginning parameters
  letters_left <- LETTERS
  guess_count <- 0
  guess_history <- data.frame(Guess = rep(NA, guess_count)) # makes a blank data frame full of NAs the length of number of guesses
  feedback_history <- data.frame(Feedback = rep(NA, guess_count))
  
  while (guess_count < num_guesses) {

  # print letters left
  print(paste(c("Letters left: ", letters_left), collapse = " "))
  
  # prompt player for a guess
  guess_count <- guess_count + 1
  guess <- toupper(readline(prompt = "Enter your guess, then press <enter>:"))
  # check if correct word length
  if (nchar(guess) != length(answer)) {
    print(paste0("⚠️ Your guess must have ", length(answer), " letters. Try again."))
    guess_count <- guess_count - 1
    next
  }
  # check if valid word
  if (!(guess %in% valid_guesses)){
    print(paste0("⚠️ Your guess is NOT a valid word. Try again."))
    guess_count <- guess_count - 1
    next
  }
  guess <- str_split_1(guess, pattern = "")
  
  # evaluate guess
  feedback <- evaluate_guess(guess, answer)
  
  # update list of letters not yet guessed
  wrong_letters <- guess[!(guess %in% answer)] # only remove letters that are not in the word at all
  letters_left <- setdiff(letters_left, wrong_letters)
  
  # update history
  guess_history[guess_count,] <- paste(guess, collapse = " ")
  feedback_history[guess_count,] <- paste(feedback, collapse = " ")
  
  # check if puzzle was solved; if not, prompt player for another guess
  # print results
  history <- cbind(guess_history, feedback_history)
  print(history)
  # if correct word is guessed
  if (all(feedback == "*")) {
    answer <- paste(answer, collapse = "")
    print(paste0("YOU WON! 🎉 The correct answer was: ", answer), quote = FALSE)
    print(paste0("Congratulations, you solved the Wordle in ", guess_count, " guesses!"), quote = FALSE)
    return()
  }
  }
  # if player runs out of guesses
  answer <- paste(answer, collapse = "")
  print(paste0("Sorry, you lost! 😭 The correct answer was: ", answer))
}
```

# To play a round of Wordle, use the functions below!

solution <- pick_solution(solution_list, word_length = "")  
game <- play_wordle(solution, valid_list, num_guess = "")

## For example:
```{r}
#| eval: false
solution <- pick_solution(solution_list, word_length = 5)
game <- play_wordle(solution, valid_list, num_guesses = 6)
```

